---
title: "Homework 6"
author: "Xinyin Miao (xm2356)"
date: "2025-11-22"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(janitor)   
library(broom)     
library(purrr)  
```

# Problem 1

```{r}
homicides <- readr::read_csv("data/homicide-data.csv") |>
  clean_names() |>
  mutate(
    city_state = str_c(city, ", ", state),
    solved = if_else(disposition == "Closed by arrest", 1L, 0L),
    victim_sex  = na_if(victim_sex,  "Unknown"),
    victim_race = na_if(victim_race, "Unknown"),
    victim_age = ifelse(victim_age %in% c("Unknown", ""), NA, victim_age),
    victim_age = as.numeric(victim_age)
    ) |>
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
         victim_race %in% c("White", "Black")
         ) |>
  mutate(
    victim_sex  = factor(victim_sex,  levels = c("Female", "Male")),
    victim_race = factor(victim_race, levels = c("White", "Black"))
  )

```

## For Baltimore,MD
```{r}
baltimore <-
  homicides |>
  filter(city_state == "Baltimore, MD")

baltimore_glm <-
  glm(
    solved ~ victim_age + victim_sex + victim_race,
    data   = baltimore,
    family = binomial()
  )

```

```{r}
baltimore_results <-
  baltimore_glm |>
  tidy(
    conf.int     = TRUE,
    exponentiate = TRUE 
  )

baltimore_male_female_or <-
  baltimore_results |>
  filter(term == "victim_sexMale") |>
  select(term, estimate, conf.low, conf.high)

baltimore_male_female_or |> 
  knitr::kable(digits = 4)

# the estimate and CI of the adjusted odds ratio for solving homicides comparing male to female victims keeping all other variables fixed

```

## For each of the cities
```{r}
city_results <-
  homicides |>
  nest(data = -city_state) |>
  mutate(
    model = map(data, \(df) glm( solved ~ victim_age + victim_sex + victim_race,  data   = df, family = binomial())
    ),
    tidied = map(model, \(m) tidy(m,
        conf.int     = TRUE,
        exponentiate = TRUE
      )
    )
  ) |>
  select(city_state, tidied) |>
  unnest(tidied) |>
  filter(term == "victim_sexMale") |> 
  select(city_state, estimate, conf.low, conf.high) |> 
  arrange(desc(estimate))

city_results |> 
  knitr::kable(digits = 4)

```

## Plot
```{r city-or-plot, fig.width=8, fig.height=7}
city_results_plot <-
  city_results |>
  mutate(
    city_state = forcats::fct_reorder(city_state, estimate)
  )

city_results_plot |>
  ggplot(aes(x = estimate, y = city_state)) +
  geom_point() +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = 0) +
  geom_vline( xintercept = 1, linetype   = "dashed") +
  labs(
    x     = "Adjusted OR of solving homicide (male vs female victims)",
    y     = "City",
    title = "Adjusted odds ratios for male vs female victims by city"
  ) +
  theme_minimal()

```

The plot shows substantial variation across cities in the adjusted odds of solving homicides for male versus female victims. Although many cities have point estimates below 1—suggesting a tendency toward lower clearance odds for male victims—the confidence intervals for nearly all cities are wide and typically cross 1. Because this uncertainty is large, the apparent pattern is not statistically reliable, and we cannot confidently conclude that cases involving female victims are more likely to be solved.

# Problem 2

```{r, message=FALSE, warning=FALSE}
library(p8105.datasets)
data("weather_df")
library(modelr)
```

## bootstrap fit
```{r}
set.seed(123)

boot_results <-
  weather_df |>
  drop_na(tmax, tmin, prcp) |>
  bootstrap(n = 5000) |>
  mutate(
    model  = map(strap, ~ lm(tmax ~ tmin + prcp, data = .x)),
    glance = map(model, broom::glance),
    tidy   = map(model, broom::tidy),
    
    r_sq = map_dbl(glance, "r.squared"),

    beta_tmin = map_dbl(tidy, ~ .x |>
                          filter(term == "tmin") |>
                          pull(estimate)),
    beta_prcp = map_dbl(tidy, ~ .x |>
                          filter(term == "prcp") |>
                          pull(estimate)),

    beta_ratio = beta_tmin / beta_prcp
  )

```

## the plot of $\hat{r^2}$
```{r}
boot_results |>
  ggplot(aes(x = r_sq)) +
  geom_histogram(bins = 30) +
  labs(
    x = expression(hat(r)^2),
    y = "Count",
    title = expression("Bootstrap distribution of " * hat(r)^2)
  ) +
  theme_bw()

```

The bootstrap distribution of $\hat{r^2}$ is approximately symmetric and bell-shaped, close to a normal distribution. 

It is centered around roughly 0.94, indicating that the fitted model consistently explains about 94% of the variability in tmax across bootstrap samples and the linear association between tmax and the predictors (tmin, prcp) is consistently strong.

It also shows moderate spread, with most values falling between 0.935 and 0.945. No extreme outliers—the tails taper smoothly on both ends.



## the plot of β1 / β2 
```{r}
boot_results |>
  ggplot(aes(x = beta_ratio)) +
  geom_histogram(bins = 30) +
  labs(
    x = expression(hat(beta)[1] / hat(beta)[2]),
    y = "Count",
    title = expression("Bootstrap distribution of " * hat(beta)[1] / hat(beta)[2])
  ) +
  theme_bw()
```

The ratio $\frac{\hat{\beta_1}}{\hat{\beta_2}}$ compares the effect of minimum temperature relative to the effect of precipitation. 

The distribution is left-skewed with a long tail extending toward more negative values.

The ratio shows substantial variability, mainly because the precipitation coefficient is small and unstable across bootstrap samples.

The long left tail (–300 to –400) indicates that small changes in the prcp coefficient can produce relatively large negative ratios.

## the 2.5% and 97.5% quantiles
```{r}
boot_results |>
  summarize(
    r2_lower      = quantile(r_sq, 0.025),
    r2_upper      = quantile(r_sq, 0.975),
    ratio_lower   = quantile(beta_ratio, 0.025),
    ratio_upper   = quantile(beta_ratio, 0.975)
  ) |> 
  knitr::kable(digits = 4, caption = "95% Bootstrap Confidence Intervals")

```



















